\textbf{A framework for effective application of machine learning to microbiome-based classification problems}

### 2020-05-14 submission/manuscript.Rmd
new 2 figures

| **Model** | **Description** | **Linearity** |
|-----------|-----------------|---------------|
|  Logistic regression |  A predictive regression analysis when the dependent variable is binary | Linear  |
|  SVM with linear kernel | A classifier that is defined by an optimal linear separating hyperplane that discriminates between labels  | Linear  |
|  SVM with radial basis kernel | A classifier that is defined by an optimal non-linear separating hyperplane that discriminates between labels  | Non-linear  |
|  Decision tree | A classifier that sorts samples down from the root to the leaf node where an attribute is tested to discriminate between labels  | Non-linear  |
|  Random forest | A classifier that is an ensemble of decision trees that grows randomly with subsampled data  | Non-linear |
|  Gradient Boosted Trees (XGBoost) | A classifier that is an ensembe of decision trees that grows greedily  | Non-linear  |

>>>>

| **Model** | **Description** | **Linearity** |
|--------------------|--------------------------------------------|---------------|
| Logistic regression | A predictive regression analysis when the dependent variable is binary |  Linear  |
| SVM with linear kernel  |  A classifier that is defined by an optimal linear separating hyperplane that discriminates between labels  | Linear  |
| SVM with radial basis kernel  | A classifier that is defined by an optimal non-linear separating hyperplane that discriminates between labels | Non-linear  |
| Decision tree | A classifier that sorts samples down from the root to the leaf node where an attribute is tested to discriminate between labels | Non-linear  |
| Random forest | A classifier that is an ensemble of decision trees that grows randomly with subsampled data  | Non-linear  |
| Gradient Boosted Trees (XGBoost)  | A classifier that is an ensembe of decision trees that grows greedily | Non-linear  |

### 2020-05-14 submission/manuscript.tex
new 2 figures

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={A framework for effective application of machine learning to microbiome-based classification problems},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

Begüm D. Topçuoğlu\({^1}\), Nicholas A. Lesniak\({^1}\), Mack
Ruffin\({^3}\), Jenna Wiens\({^2}\), Patrick D.
Schloss\textsuperscript{1\(\dagger\)}

@@

\captionsetup{labelformat=empty}
\captionof{table}{\textbf{Table 1.} Characteristics of the machine learning models in our comparative study.}
\small

\begin{tabular}{|l|l|l|}
\hline

\rowcolor{lightgray}
\textbf{Model} & \textbf{Description} & \textbf{Linearity} \\ \hline

\makecell[l]{Logistic \\regression} & \makecell[l]{A predictive regression analysis when the dependent \\variable is binary.} & Linear \\ \hline

\makecell[l]{SVM with \\linear kernel} & \makecell[l]{A classifier that is defined by an optimal linear \\separating hyperplane that discriminates between labels.} & Linear  \\ \hline

\makecell[l]{SVM with \\radial basis kernel} & \makecell[l]{A classifier that is defined by an optimal non-linear \\separating hyperplane that discriminates between labels.} & Non-linear \\ \hline

\makecell[l]{Decision tree} & \makecell[l]{A classifier that sorts samples down from the
root to the \\leaf node where an attribute is tested to discriminate \\between labels.} & Non-linear \\ \hline

Random forest & \makecell[l]{A classifier that is an ensemble of decision trees \\ that grows randomly with subsampled data.} & Non-linear \\ \hline

\makecell[l]{Gradient Boosted Trees \\ (XGBoost)} & \makecell[l]{A classifier that is an ensembe of decision trees \\ that grows greedily.} & Non-linear \\ \hline

\end{tabular}\newpage

>>>>

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={A framework for effective application of machine learning to microbiome-based classification problems},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

Begüm D. Topçuoğlu\({^1}\), Nicholas A. Lesniak\({^1}\), Mack
Ruffin\({^3}\), Jenna Wiens\textsuperscript{2\(\dagger\)}, Patrick D.
Schloss\textsuperscript{1\(\dagger\)}

@@

\textbf{Table 1.} Characteristics of the machine learning models in our
comparative study.

\begin{longtable}[]{@{}lll@{}}
\toprule
\begin{minipage}[b]{0.23\columnwidth}\raggedright\strut
\textbf{Model}\strut
\end{minipage} & \begin{minipage}[b]{0.50\columnwidth}\raggedright\strut
\textbf{Description}\strut
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\raggedright\strut
\textbf{Linearity}\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Logistic regression\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
A predictive regression analysis when the dependent variable is
binary\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Linear\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
SVM with linear kernel\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
A classifier that is defined by an optimal linear separating hyperplane
that discriminates between labels\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Linear\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
SVM with radial basis kernel\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
A classifier that is defined by an optimal non-linear separating
hyperplane that discriminates between labels\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Non-linear\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Decision tree\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
A classifier that sorts samples down from the root to the leaf node
where an attribute is tested to discriminate between labels\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Non-linear\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Random forest\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
A classifier that is an ensemble of decision trees that grows randomly
with subsampled data\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Non-linear\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Gradient Boosted Trees (XGBoost)\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
A classifier that is an ensembe of decision trees that grows
greedily\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Non-linear\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}
